### 2020-01-24 abstract.tex
initial submission

Many programming languages offer multiple static analysis tools that offer to detect faults in code witthout executing it.  Understanding the strengths and weaknesses of tools, and performing direct comparisons of their effectiveness is diffcult; it usually involves either manual examination of differing warnings on real code, or the bias-prone construction of artificial test cases.  In practice, comparisons tend to be limited to superficial, anecdotal discussions in the informal literature (e.g., blog posts by software developers), or purely research-community-oriented evaluations made by the authors of new tools seeking to publish their results.  This paper proposes a novel automated approach to comparing static analysis tools, based on producing \emph{mutants} of real code, and comparing mutation detection rates for tools to their warning rates on the original code.  In addition to making tool differences quantitatively observable without extensive manual effort, this approach offers a new way to detect and fix omissions in a static analysis tool's set of detectors.  We present an extensive comparison of three well-known Solidity smart contract static analysis tools, and show how using an automatic prioritization of our results allowed us to add three new detectors to the best of the tools.  We also evaluate popular Java and Python static analysis tools and discuss their strengths and weaknesses.

>>>>

Many programming languages offer multiple static analysis tools that offer to detect faults in code without executing it.  Understanding the strengths and weaknesses of tools, and performing direct comparisons of their effectiveness is difficult; it usually involves either manual examination of differing warnings on real code, or the bias-prone construction of artificial test cases.  In practice, comparisons tend to be limited to superficial, anecdotal discussions in the informal literature (e.g., blog posts by software developers), or purely research-community-oriented evaluations made by the authors of new tools seeking to publish their results.  This paper proposes a novel automated approach to comparing static analysis tools, based on producing \emph{mutants} of real code, and comparing mutation detection rates for tools to their warning rates on the original code.  In addition to making tool differences quantitatively observable without extensive manual effort, this approach offers a new way to detect and fix omissions in a static analysis tool's set of detectors.  We present an extensive comparison of three well-known Solidity smart contract static analysis tools, and show how using an automatic prioritization of our results allowed us to add three new detectors to the best of the tools.  We also evaluate popular Java and Python static analysis tools and discuss their strengths and weaknesses.

### 2020-01-24 conclusions.tex
more python
initial submission

In this paper, we showed that program mutants can be used as a proxy
for real faults, to compare (and motivate improvements to) static
analysis tools.  Mutants are attractive in that changing a mostly
correct program usually introduces a bug into it; this is the basis of
mutation testing, after all, and a large body of work supports the
claim that at least 60-70\% of mutants are fault-inducing.   This
means we can assume mutants are faulty, and escape the
ground-truth/false positive problem that makes comparing static
analysis tools so labor-intensive.  It cannot help with precision
problems, directly, but combined with finding counts for un-mutated
code, our approach can identify tools that detect mutants well,
adjusted for their general tendency to flag any code as faulty.  We
evaluated 9 popular static analysis tools, for Solidity smart contracts, Java,
and Python, and offer advice to users of these tools.  We were also
able to use our methods, plus a novel mutant prioritization scheme, to
identify three useful new detectors for the Slither smart contract
analyzer.

As future work, we would like to further validate our approach and
improve our admittedly \emph{ad hoc} mutant distance metric.  Allowing
user feedback \cite{EndUserMistake,OnlyOracle}, or applying metric
learning methods \cite{kulis2012metric} (particularly unsupervised
learning \cite{scholkopf1998nonlinear,tipping1999probabilistic}) are
the most obvious and interesting possibilites for a better metric.
Finally, mutant prioritization should be applicable to improving
software testing, as well \cite{groce2018verified}.

>>>>

In this paper, we showed that program mutants can be used as a proxy
for real faults, to compare (and motivate improvements to) static
analysis tools.  Mutants are attractive in that changing a mostly
correct program usually introduces a bug into it; this is the basis of
mutation testing, after all, and a large body of work supports the
claim that at least 60-70\% of mutants are fault-inducing.   This
means we can assume mutants are faulty, and escape the
ground-truth/false positive problem that makes comparing static
analysis tools so labor-intensive.  Our approach cannot identify precision
problems, directly, but combined with finding counts for un-mutated
code, our approach can identify tools that detect mutants well,
adjusted for their general tendency to flag any code as faulty, and
thus identify imprecise tools.  We
evaluated 9 popular static analysis tools, for Solidity smart contracts, Java,
and Python, and offer advice to users of these tools.  Our mutation
results strongly confirm the wisdom of using multiple tools; with the
exception of one Java tool, all tools we investigated uniquely
detected over 1,000 mutants, and for Java and Python there were no
mutants detected by all tools. For Solidity,
academic research evaluations of the tools generally agreed strongly
with our conclusions, but lacked the detail mutant analysis contributed.
We were also able to use our methods, plus a novel mutant prioritization scheme, to
identify three useful new detectors for the Slither smart contract
analyzer.

As future work, we would like to further validate our approach and
improve our admittedly \emph{ad hoc} mutant distance metric.  Allowing
user feedback \cite{EndUserMistake,OnlyOracle}, or applying metric
learning methods \cite{kulis2012metric} (particularly unsupervised
learning \cite{scholkopf1998nonlinear,tipping1999probabilistic}) are
the most obvious and interesting possibilities for a better metric.
Finally, mutant prioritization should be applicable to improving
software testing, as well \cite{groce2018verified}.
